---
title: "Data Preparing"
author: "JIN QINHAO"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    page-layout: full
execute:
  warning: false
  message: false
---

# **Visualization of classification trees and random forest analysis**

# 1Data Preparations

## 1.1Install and launch the required R packages

```{r}
pacman::p_load(
  tidyverse,
  readxl,
  lubridate,
  writexl
)
```

## 1.2Importing Data

```{r}
base_dir <- "D:/GitHub/Qinhao/ISSS608-VAA/Take-home-exercise/Take-Home-Exercise2"
raw_path <- file.path(base_dir, "data", "Name your insight (2).xlsx")
processed_dir <- file.path(base_dir, "data")
artifact_export_enabled <- identical(
  tolower(Sys.getenv("WRITE_TAKE_HOME_EX2_ARTIFACTS", "false")),
  "true"
)

export_note <- function(path_label) {
  message(
    "Artifact export skipped during render. ",
    "Set WRITE_TAKE_HOME_EX2_ARTIFACTS=true to regenerate tracked files under: ",
    path_label
  )
}

if (artifact_export_enabled) {
  dir.create(processed_dir, recursive = TRUE, showWarnings = FALSE)
}
```

## 1.3Read the original data

```{r}
col_names <- c(
  "date",
  "avg_stay_monthly",
  "hotel_occ",
  "spend_per_capita",
  "tourism_receipts",
  "avg_stay_annual",
  "visitor_arrivals",
  "visitor_arrivals_china"
)

raw_tbl <- readxl::read_excel(
  path = raw_path,
  sheet = "My Series",
  skip = 29,
  col_names = FALSE
) %>%
  setNames(col_names) %>%
  mutate(
    date = as.Date(date),
    across(-date, as.numeric)
  ) %>%
  arrange(date)
```

## 1.4View the original data structure

```{r}
glimpse(raw_tbl)
colSums(is.na(raw_tbl))
```

## 1.5Retain the valid data suitable for monthly analysis

```{r}
monthly_clean <- raw_tbl %>%
  filter(
    !is.na(date),
    !is.na(avg_stay_monthly),
    !is.na(hotel_occ),
    !is.na(visitor_arrivals),
    !is.na(visitor_arrivals_china)
  ) %>%
  mutate(
    year = lubridate::year(date),
    month = lubridate::month(date),
    quarter = lubridate::quarter(date),
    period = case_when(
      date <= as.Date("2020-01-01") ~ "pre_covid",
      date <= as.Date("2021-12-01") ~ "covid_shock",
      TRUE ~ "recovery"
    ),
    china_share = visitor_arrivals_china / visitor_arrivals,
    china_share_pct = china_share * 100
  )
```

## 1.6Handling extreme values

```{r}
stay_cap <- quantile(monthly_clean$avg_stay_monthly, probs = 0.95, na.rm = TRUE)

monthly_clean <- monthly_clean %>%
  mutate(
    avg_stay_monthly_capped = pmin(avg_stay_monthly, stay_cap),
    visitor_arrivals_millions = visitor_arrivals / 1000000,
    visitor_arrivals_china_thousands = visitor_arrivals_china / 1000
  )
```

## 1.7Generate analysis fields

```{r}
occ_cuts <- quantile(monthly_clean$hotel_occ, probs = c(1/3, 2/3), na.rm = TRUE)

analysis_ready <- monthly_clean %>%
  mutate(
    cluster_z_visitor_arrivals = as.numeric(scale(visitor_arrivals)),
    cluster_z_china_share = as.numeric(scale(china_share)),
    cluster_z_hotel_occ = as.numeric(scale(hotel_occ)),
    cluster_z_avg_stay_monthly_capped = as.numeric(scale(avg_stay_monthly_capped)),
    hotel_occ_level_tertile = case_when(
      hotel_occ <= occ_cuts[[1]] ~ "low",
      hotel_occ <= occ_cuts[[2]] ~ "medium",
      TRUE ~ "high"
    ),
    hotel_occ_level_business = case_when(
      hotel_occ < 70 ~ "low",
      hotel_occ <= 85 ~ "medium",
      TRUE ~ "high"
    )
  ) %>%
  mutate(
    hotel_occ_level_tertile = factor(hotel_occ_level_tertile, levels = c("low", "medium", "high")),
    hotel_occ_level_business = factor(hotel_occ_level_business, levels = c("low", "medium", "high"))
  ) %>%
  mutate(
    dataset_split = if_else(row_number() <= floor(n() * 0.8), "train", "test")
  )

```

## 1.8Generate data specifically for classification tree construction

```{r}
decision_tree_ready <- analysis_ready %>%
  select(
    date, year, month, quarter, period,
    visitor_arrivals, visitor_arrivals_china,
    china_share, china_share_pct,
    hotel_occ, avg_stay_monthly, avg_stay_monthly_capped,
    hotel_occ_level_tertile, hotel_occ_level_business,
    dataset_split
  )


```

## 1.9Export the cleaned data

```{r}
if (artifact_export_enabled) {
  readr::write_csv(monthly_clean, file.path(processed_dir, "tourism_monthly_clean.csv"))
  readr::write_csv(decision_tree_ready, file.path(processed_dir, "tourism_decision_tree_ready.csv"))
  readr::write_csv(analysis_ready, file.path(processed_dir, "tourism_four_part_analysis_ready.csv"))

  writexl::write_xlsx(
    list(
      monthly_clean = monthly_clean,
      decision_tree_ready = decision_tree_ready,
      analysis_ready_all = analysis_ready
    ),
    path = file.path(processed_dir, "tourism_four_part_analysis_ready.xlsx")
  )
} else {
  export_note(processed_dir)
}


```

## 1.10Output summary

```{r}
cat("Data preparation complete.\n")
cat("Rows in monthly_clean:", nrow(monthly_clean), "\n")
cat("Rows in analysis_ready:", nrow(analysis_ready), "\n")
cat("Date range:", min(analysis_ready$date), "to", max(analysis_ready$date), "\n")
cat("95% cap for avg_stay_monthly:", round(stay_cap, 6), "\n")
cat("Hotel occupancy cut points:", round(occ_cuts[[1]], 6), "and", round(occ_cuts[[2]], 6), "\n")

```
