---
title: "Visualization of  Random Forest "
author: "JIN QINHAO"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    page-layout: full
execute:
  warning: false
  message: false
---

# **Visualization of random forest analysis**

## 1.1Install and launch the required R packages

```{r}
pacman::p_load(
  tidyverse,
  rpart,
  rpart.plot,
  caret,
  visNetwork,
  htmlwidgets,
  sparkline
)
```

```{r}
base_dir <- "D:/GitHub/Qinhao/ISSS608-VAA/Take-home-exercise/Take-Home-Exercise2"
data_path <- file.path(base_dir, "data", "tourism_decision_tree_ready.csv")
output_dir <- file.path(base_dir, "outputs")

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

```

## 1.2Reading data

```{r}
df <- readr::read_csv(data_path, show_col_types = FALSE) %>%
  mutate(
    date = as.Date(date),
    month = factor(
      month,
      levels = 1:12,
      labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
    ),
    quarter = factor(quarter),
    period = factor(period, levels = c("pre_covid", "covid_shock", "recovery")),
    dataset_split = factor(dataset_split, levels = c("train", "test")),
    hotel_occ_level_tertile = factor(
      hotel_occ_level_tertile,
      levels = c("low", "medium", "high")
    ),

    # 为了让树图上的阈值更易读，创建展示型变量
    arrivals_million = visitor_arrivals / 1000000,
    china_share_pct = china_share * 100,
    stay_days = avg_stay_monthly_capped
  )

glimpse(df)


```

## 1.3Split the dataset into training and testing sets

```{r}
train_df <- df %>% filter(dataset_split == "train")
test_df  <- df %>% filter(dataset_split == "test")

cat("Train rows:", nrow(train_df), "\n")
cat("Test rows:", nrow(test_df), "\n")

```

# 2Random Forest Analysis Hyperparameter tuning

```{r}
set.seed(1234)

ctrl <- trainControl(
  method = "repeatedcv",
  number = 5,
  repeats = 3,
  classProbs = TRUE
)

grid <- expand.grid(
  mtry = c(2, 3, 4),
  splitrule = "gini",
  min.node.size = c(1, 3, 5)
)

rf_model <- caret::train(
  hotel_occ_level_tertile ~ arrivals_million + china_share_pct + stay_days + month,
  data = train_df,
  method = "ranger",
  trControl = ctrl,
  tuneGrid = grid,
  num.trees = 300,
  importance = "impurity",
  metric = "Accuracy"
)

print(rf_model)
print(rf_model$bestTune)
```

## 2.1Predict on test set

```{r}
pred_class <- predict(rf_model, newdata = test_df)
pred_prob  <- predict(rf_model, newdata = test_df, type = "prob")

pred_class <- factor(pred_class, levels = c("low", "medium", "high"))
test_df$hotel_occ_level_tertile <- factor(
  test_df$hotel_occ_level_tertile,
  levels = c("low", "medium", "high")
)
```

## 2.2Model evaluation

```{r}
cm <- caret::confusionMatrix(
  data = pred_class,
  reference = test_df$hotel_occ_level_tertile
)

print(cm)

metrics_tbl <- tibble(
  metric = c("accuracy", "kappa"),
  value = c(cm$overall[["Accuracy"]], cm$overall[["Kappa"]])
)

print(metrics_tbl)
```

## 2.3Variable importance

```{r}
importance_tbl <- varImp(rf_model)$importance %>%
  tibble::rownames_to_column("variable") %>%
  arrange(desc(Overall))

print(importance_tbl)
```

## 2.4Confusion matrix heatmap

```{r}
conf_tbl <- as.data.frame(cm$table)

conf_plot <- ggplot(conf_tbl, aes(x = Reference, y = Prediction, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = Freq), size = 5) +
  scale_fill_gradient(low = "#f7fbff", high = "#08519c") +
  labs(
    title = "Random Forest Confusion Matrix",
    x = "Actual class",
    y = "Predicted class",
    fill = "Count"
  ) +
  theme_minimal(base_size = 13)

print(conf_plot)

ggsave(
  filename = file.path(output_dir, "random_forest_confusion_heatmap.png"),
  plot = conf_plot,
  width = 8,
  height = 6,
  dpi = 200
)
```

## 2.5Variable importance plot

```{r}
importance_plot <- importance_tbl %>%
  slice_max(order_by = Overall, n = 10) %>%
  ggplot(aes(x = reorder(variable, Overall), y = Overall)) +
  geom_col(fill = "#2c7fb8") +
  coord_flip() +
  labs(
    title = "Random Forest Variable Importance",
    x = NULL,
    y = "Importance"
  ) +
  theme_minimal(base_size = 13)

print(importance_plot)

ggsave(
  filename = file.path(output_dir, "random_forest_importance.png"),
  plot = importance_plot,
  width = 8,
  height = 6,
  dpi = 200
)

```

## 2.6OOB accuracy vs number of trees

```{r}
tree_grid <- seq(50, 300, by = 25)

oob_tbl <- purrr::map_dfr(tree_grid, function(nt) {
  mod <- ranger::ranger(
    hotel_occ_level_tertile ~ arrivals_million + china_share_pct + stay_days + month,
    data = train_df,
    num.trees = nt,
    mtry = rf_model$bestTune$mtry,
    min.node.size = rf_model$bestTune$min.node.size,
    splitrule = as.character(rf_model$bestTune$splitrule),
    importance = "impurity",
    seed = 1234
  )

  tibble(
    num_trees = nt,
    oob_error = mod$prediction.error,
    oob_accuracy = 1 - mod$prediction.error
  )
})

print(oob_tbl)

oob_plot <- ggplot(oob_tbl, aes(x = num_trees, y = oob_accuracy)) +
  geom_line(color = "#238b45", linewidth = 1) +
  geom_point(color = "#238b45", size = 2) +
  labs(
    title = "OOB Accuracy vs Number of Trees",
    x = "Number of trees",
    y = "OOB accuracy"
  ) +
  theme_minimal(base_size = 13)

print(oob_plot)

ggsave(
  filename = file.path(output_dir, "random_forest_oob_accuracy.png"),
  plot = oob_plot,
  width = 8,
  height = 6,
  dpi = 200
)
```

## 2.7Actual vs predicted over time

```{r}
pred_plot_df <- test_df %>%
  transmute(
    date,
    actual = hotel_occ_level_tertile,
    predicted = pred_class
  ) %>%
  pivot_longer(
    cols = c(actual, predicted),
    names_to = "series",
    values_to = "level"
  )

pred_plot <- ggplot(pred_plot_df, aes(x = date, y = level, color = series, shape = series)) +
  geom_point(size = 3) +
  labs(
    title = "Actual vs Predicted Hotel Occupancy Level",
    x = "Date",
    y = "Occupancy level"
  ) +
  scale_color_manual(values = c(actual = "#d95f0e", predicted = "#3182bd")) +
  theme_minimal(base_size = 13)

print(pred_plot)

ggsave(
  filename = file.path(output_dir, "random_forest_actual_vs_predicted.png"),
  plot = pred_plot,
  width = 9,
  height = 5,
  dpi = 200
)
```

## 2.8Export results

```{r}
write.csv(
  metrics_tbl,
  file.path(output_dir, "random_forest_metrics.csv"),
  row.names = FALSE
)

write.csv(
  conf_tbl,
  file.path(output_dir, "random_forest_confusion_matrix.csv"),
  row.names = FALSE
)

write.csv(
  importance_tbl,
  file.path(output_dir, "random_forest_variable_importance.csv"),
  row.names = FALSE
)

write.csv(
  oob_tbl,
  file.path(output_dir, "random_forest_oob_accuracy.csv"),
  row.names = FALSE
)

prediction_tbl <- bind_cols(
  test_df %>% select(date, hotel_occ_level_tertile),
  tibble(predicted_class = pred_class),
  pred_prob
)

write.csv(
  prediction_tbl,
  file.path(output_dir, "random_forest_test_predictions.csv"),
  row.names = FALSE
)

cat("Random forest analysis complete.\n")
cat("Outputs saved to:", output_dir, "\n")
```
