---
title: "Visualization of classification trees "
author: "JIN QINHAO"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    code-summary: "Show the code"
    page-layout: full
execute:
  warning: false
  message: false
---

# **Visualization of classification trees**

# 1Classification tree analysis

## 1.1Install and launch the required R packages

```{r}
pacman::p_load(
  tidyverse,
  rpart,
  rpart.plot,
  caret,
  visNetwork,
  htmlwidgets,
  sparkline
)
```

```{r}
base_dir <- "D:/GitHub/Qinhao/ISSS608-VAA/Take-home-exercise/Take-Home-Exercise2"
data_path <- file.path(base_dir, "data", "tourism_decision_tree_ready.csv")
output_dir <- file.path(base_dir, "outputs")
artifact_export_enabled <- identical(
  tolower(Sys.getenv("WRITE_TAKE_HOME_EX2_ARTIFACTS", "false")),
  "true"
)

export_note <- function(path_label) {
  message(
    "Artifact export skipped during render. ",
    "Set WRITE_TAKE_HOME_EX2_ARTIFACTS=true to regenerate tracked files under: ",
    path_label
  )
}

if (artifact_export_enabled) {
  dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
}

```

## 1.2Reading data

```{r}
df <- readr::read_csv(data_path, show_col_types = FALSE) %>%
  mutate(
    date = as.Date(date),
    month = factor(
      month,
      levels = 1:12,
      labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                 "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
    ),
    quarter = factor(quarter),
    period = factor(period, levels = c("pre_covid", "covid_shock", "recovery")),
    dataset_split = factor(dataset_split, levels = c("train", "test")),
    hotel_occ_level_tertile = factor(
      hotel_occ_level_tertile,
      levels = c("low", "medium", "high")
    ),

    # 为了让树图上的阈值更易读，创建展示型变量
    arrivals_million = visitor_arrivals / 1000000,
    china_share_pct = china_share * 100,
    stay_days = avg_stay_monthly_capped
  )

glimpse(df)


```

## 1.3Split the dataset into training and testing sets

```{r}
train_df <- df %>% filter(dataset_split == "train")
test_df  <- df %>% filter(dataset_split == "test")

cat("Train rows:", nrow(train_df), "\n")
cat("Test rows:", nrow(test_df), "\n")

```

## 1.4Build classification tree

Target variable: Hotel occupancy rate level (low / medium / high)

Independent variables: \# arrivals_million = Total inbound tourists (in millions)

china_share_pct = Proportion of Chinese tourists (percentage)

stay_days = Truncated average length of stay

month = Month (seasonality)

```{r}
fit_tree <- rpart::rpart(
  hotel_occ_level_tertile ~ arrivals_million + china_share_pct + stay_days + month,
  data = train_df,
  method = "class",
  parms = list(split = "gini"),
  control = rpart::rpart.control(
    cp = 0.005,
    maxdepth = 4,
    minsplit = 10,
    minbucket = 5,
    xval = 10
  )
)

```

## 1.5Check the complexity parameter table.

```{r}
printcp(fit_tree)
plotcp(fit_tree)
```

## 1.6Prune the tree

```{r}

bestcp <- fit_tree$cptable[which.min(fit_tree$cptable[, "xerror"]), "CP"]
pruned_tree <- prune(fit_tree, cp = bestcp)

cat("Best cp:", bestcp, "\n")
print(pruned_tree)
```

## 1.7Interactive tree visualization

```{r}
tree_vis <- visNetwork::visTree(
  pruned_tree,
  data = as.data.frame(train_df),
  main = "Decision Tree for Hotel Occupancy Level",
  submain = "Target: low / medium / high",
  width = "100%",
  height = "700px",
  nodesFontSize = 16,
  edgesFontSize = 14,
  fallenLeaves = TRUE,
  legend = TRUE,
  nodesPopSize = TRUE,
  minNodeSize = 14,
  maxNodeSize = 32,
  highlightNearest = list(enabled = TRUE, hover = TRUE),
  export = TRUE
)

tree_vis

```

```{r}
if (artifact_export_enabled) {
  htmlwidgets::saveWidget(
    tree_vis,
    file.path(output_dir, "decision_tree_visual.html"),
    selfcontained = FALSE
  )
} else {
  export_note(output_dir)
}

```

## 1.8Static tree visualization for report

```{r}
if (artifact_export_enabled) {
  png(
    filename = file.path(output_dir, "decision_tree_static.png"),
    width = 1800,
    height = 1200,
    res = 180
  )

  rpart.plot::rpart.plot(
    pruned_tree,
    type = 2,
    extra = 104,
    under = TRUE,
    fallen.leaves = TRUE,
    faclen = 0,
    tweak = 1.15,
    main = "Decision Tree for Hotel Occupancy Level"
  )

  dev.off()
} else {
  export_note(output_dir)
}

```

## 1.9Predict on test set

```{r}
pred_class <- predict(pruned_tree, newdata = test_df, type = "class")
pred_prob  <- predict(pruned_tree, newdata = test_df, type = "prob") %>%
  as.data.frame()

# 统一 levels，避免 confusionMatrix 报错
pred_class <- factor(pred_class, levels = c("low", "medium", "high"))
test_df$hotel_occ_level_tertile <- factor(
  test_df$hotel_occ_level_tertile,
  levels = c("low", "medium", "high")
)

```

## 1.10Evaluate model performance

```{r}
cm <- caret::confusionMatrix(
  data = pred_class,
  reference = test_df$hotel_occ_level_tertile
)

print(cm)

metrics_tbl <- tibble(
  metric = c("accuracy", "kappa"),
  value = c(cm$overall[["Accuracy"]], cm$overall[["Kappa"]])
)

print(metrics_tbl)

```

## 1.11Variable importance

```{r}
importance_tbl <- tibble(
  variable = names(pruned_tree$variable.importance),
  importance = as.numeric(pruned_tree$variable.importance)
) %>%
  arrange(desc(importance))

print(importance_tbl)
```

## 1.12Export results

```{r}
prediction_tbl <- bind_cols(
  test_df %>% select(date, hotel_occ, hotel_occ_level_tertile),
  tibble(predicted_class = pred_class),
  pred_prob
)

if (artifact_export_enabled) {
  write.csv(
    as.data.frame(cm$table),
    file.path(output_dir, "decision_tree_confusion_matrix.csv"),
    row.names = FALSE
  )

  write.csv(
    metrics_tbl,
    file.path(output_dir, "decision_tree_metrics.csv"),
    row.names = FALSE
  )

  write.csv(
    importance_tbl,
    file.path(output_dir, "decision_tree_variable_importance.csv"),
    row.names = FALSE
  )

  write.csv(
    prediction_tbl,
    file.path(output_dir, "decision_tree_test_predictions.csv"),
    row.names = FALSE
  )

  cat("Decision tree analysis complete.\n")
  cat("Outputs saved to:", output_dir, "\n")
} else {
  cat("Decision tree analysis complete.\n")
  export_note(output_dir)
}
```
