---
title: "Hands-on Exercise 1"
author: "JIN QINHAO"
---

# Getting Started

## Install and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

#Essential Grammatical Elements in ggplot2:geom

## Bar Chart

```{r}
#| echo: false
ggplot(data = exam_data,
       aes(x = RACE)) + 
  geom_bar()
```

## Dot Plot

```{r}
#| echo: false
ggplot(data=exam_data,
       aes(x = MATHS)) +
  geom_dotplot(dotsize = 0.5)
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  
  geom_dotplot(binwidth=2.5,
               dotsize= 0.5) +
  scale_y_continuous(NULL,
                      breaks = NULL)
```

## Histogram

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram()
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20,
                 color = "black",
                 fill = "lightblue")
```

### Modify Via aes changes

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           fill = GENDER)) +
  geom_histogram(bins = 20,
                color = "gray")
```

## Density Graph

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_density()
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           colours = GENDER)) +
  geom_density()
```

## Box Plot

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = GENDER)) +
  geom_boxplot()
```

### Box Plot With Notches (Show Median)

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = GENDER)) + 
  geom_boxplot(notch = TRUE)
```

## Violin Graph (geom_violin())

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = GENDER)) +
  geom_violin()
```

## Point Graph (geom_point())

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) +
  geom_point()
```

## Combined Graph (Box plot + Point)

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = GENDER)) +
  geom_boxplot() +
  geom_point(position = "jitter",
             size = 0.5)

```

# Working With Stats()

## stat_summary() method

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",
               fun = "mean",
               colour = "red",
               size = 4)


```

## geom() method

```{r}
ggplot(data = exam_data,
       aes(y = MATHS,
           x = GENDER)) +
  geom_boxplot() +
  geom_point(stat = "summary",
             fun = "mean",
             colour = "red",
             size = 4)

```

## Addming Best Fit Curve for SCatter plot

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) + 
  geom_point() + 
  geom_smooth(size = 0.5)
```

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) + 
  geom_point() + 
  geom_smooth(method = lm,
              linewidth = 0.5)
```

# Working with facet_warp()

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) +
  geom_histogram(bins = 20) +
  facet_wrap(~ CLASS)
```

## facet_grid()

```{r}
ggplot(data = exam_data,
       aes(x = MATHS)) + 
  geom_histogram(bins = 20) +
  facet_grid(~ CLASS)
```

# Using Coordinate

## coord_flip()

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) +
  geom_bar() +
  coord_flip()
```

### fixing x and y range using coord_cartesian()

```{r}
ggplot(data = exam_data,
       aes(x = MATHS,
           y = ENGLISH)) + 
  geom_point() + 
  geom_smooth(method = lm,
              size = 0.5) +
  coord_cartesian(xlim = c(0,100),
                  ylim = c(0,100))
```

# Working With Theme - \[Background Colour\] \| \[Font Size\] \| \[Grid Lines\] \| \[Label Colours\]

## theme_gray()

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) + 
  geom_bar() +
  coord_flip() +
  theme_gray()
```

## theme_classic()

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) + 
  geom_bar() +
  coord_flip() +
  theme_classic()
```

## Theme Minimal

```{r}
ggplot(data = exam_data,
       aes(x = RACE)) + 
  geom_bar() +
  coord_flip() +
  theme_minimal()
```
