---
title: "Hands-on Exercise 5d"
author: "JIN QINHAO"
---

## **Getting Started**

### **Installing and loading the packages**

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

## Importing the data

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

## **Plotting Static Parallel Coordinates**

### **Plotting a simple parallel coordinates**

```{r}
ggparcoord(data = wh, 
           columns = c(7:12))
```

### **Plotting a parallel coordinates with boxplot**

```{r, warning=FALSE, message=FALSE}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```

## **Parallel coordinates with facet**

```{r, warning=FALSE, message=FALSE}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region)
```

### **Rotating x-axis text label**

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30))
```

## **Adjusting the rotated x-axis text label**

```{r, warning=FALSE, message=FALSE}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

### **Adjusting the rotated x-axis text label**

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Multiple Parallel Coordinates Plots of World Happines Variables by Region") +
  facet_wrap(~ Region) + 
  theme(axis.text.x = element_text(angle = 30, hjust=1))
```

## **Plotting Interactive Parallel Coordinates Plot: parallelPlot method**

### **The basic plot**

```{r}
wh <- wh %>%
  select("Happiness score", c(7:12))
parallelPlot(wh,
             width = 320,
             height = 250)
```

### **Rotate axis label**

```{r}
parallelPlot(wh,
             continuousCS = "YlOrRd",
             rotateTitle = TRUE)
```

### **Parallel coordinates plot with histogram**

```{r}
histoVisibility <- rep(TRUE, ncol(wh))
parallelPlot(wh,
             rotateTitle = TRUE,
             histoVisibility = histoVisibility)
```
