---
title: "Hands-on Exercise 5c"
author: "JIN QINHAO"
---

## **Getting Started**

### **Installing and loading the packages**

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

## Importing the data

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

### **Preparing the data**

### **Transforming the data frame into a matrix**

```{r}

row.names(wh) <- wh$Country

wh1 <- dplyr::select(wh, c(3, 7:12))
wh_matrix <- data.matrix(wh)
```

### **heatmap() of R Stats**

```{r, warning=FALSE, message=FALSE}
wh_heatmap <- heatmap(wh_matrix,
                      Rowv=NA, Colv=NA)
```

to plot a cluster heatmap, we just have to use the default as shown in the code chunk below.

```{r}
wh_heatmap <- heatmap(wh_matrix)
```

### **Plotting confidence interval of point estimates**

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",
                      cexRow = 0.6, 
                      cexCol = 0.8,
                      margins = c(10, 4))
```

## **Creating Interactive Heatmap**

### **Working with heatmaply**

```{r, warning=FALSE, message=FALSE}
heatmaply(
  mtcars,
  width  = 600,
  height = 400
)
```

### 

```{r}

mat <- as.matrix(wh_matrix[, -c(1, 2, 4, 5)])
mode(mat) <- "numeric"

heatmaply(
  mat,
  width  = 600,
  height = 500
)

```

### **Data trasformation**

#### Scaling method

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column",  width  = 1200,
  height = 2000)
```

#### Normalising method

```{r, warning=FALSE, message=FALSE}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))

```

#### Normalising method

```{r}

heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

#### Percentising method

```{r}
heatmaply(percentize(wh_matrix[, -c(1, 2, 4, 5)]))
```

### **Clustering algorithm**

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```

### **Statistical approach**

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]
```

```{r}
wh_clust <- hclust(wh_d, method = "average")
num_k <- find_k(wh_clust)
plot(num_k)
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```

### **Seriation**

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "OLO")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "mean")
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none")
```

### **Working with colour palettes**

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "none",
          colors = Blues)
```

### **The finishing touch**

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,
          k_row = 5,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```
