---
title: "Hands-on Exercise 5b"
author: "JIN QINHAO"
---

# **Visual Correlation Analysis**

## **Getting Started**

## **Installing and Launching R Packages**

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

## Importing the data

```{r}
wine <- read_csv("data/wine_quality.csv")

```

## **Building Correlation Matrix: *pairs()* method**

### **Building a basic correlation matrix**

```{r}
#| fig-width: 12
#| fig-height: 12
pairs(wine[, 1:11])
```

```{r}
pairs(wine[,2:12])
```

### **Drawing the lower corner**

```{r}
pairs(wine[,2:12], upper.panel = NULL)

```

Similarly, you can display the upper half of the correlation matrix by using the code chun below.

```{r}
pairs(wine[,2:12], lower.panel = NULL)
```

### **Including with correlation coefficients**

```{r,warning=FALSE,error=FALSE}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```

## **Visualising Correlation Matrix: ggcormat():The basic plot**

```{r, warning=FALSE, message=FALSE}
#| fig-width: 10
#| fig-height: 10
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11)
```

```{r}
#| fig-width: 10
#| fig-height: 10
ggstatsplot::ggcorrmat(
  data = wine, 
  cor.vars = 1:11,
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  title    = "Correlogram for wine dataset",
  subtitle = "Four pairs are no significant at p < 0.05"
)
```

Things to learn from the code chunk above:

-   `cor.vars` argument is used to compute the correlation matrix needed to build the corrgram.

-   `ggcorrplot.args` argument provide additional (mostly aesthetic) arguments that will be passed to [`ggcorrplot::ggcorrplot`](http://www.sthda.com/english/wiki/ggcorrplot-visualization-of-a-correlation-matrix-using-ggplot2) function. The list should avoid any of the following arguments since they are already internally being used: `corr`, `method`, `p.mat`, `sig.level`, `ggtheme`, `colors`, `lab`, `pch`, `legend.title`, `digits`.

The sample sub-code chunk can be used to control specific component of the plot such as the font size of the x-axis, y-axis, and the statistical report.

```{r}
ggplot.component = list(
    theme(text=element_text(size=5),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8)))
```

## **Building multiple plots**

```{r}
#| fig-width: 14
#| fig-height: 14
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```

## **Visualising Correlation Matrix using corrplot Package**

### **Getting started with corrplot**

```{r}
#| fig-width: 14
#| fig-height: 14
library(corrplot)

wine.cor <- cor(wine[, 1:11])
corrplot(
  wine.cor,
  method = "circle",
  type   = "full",
  tl.cex = 1.5
)

```

### **Working with visual geometrics**

```{r}
#| fig-width: 14
#| fig-height: 14
corrplot(wine.cor, 
         method = "ellipse",
         type   = "full",
         tl.cex = 1.5
         ) 
```

### **Working with layout**

```{r}
#| fig-width: 14
#| 
#| fig-height: 14
corrplot(wine.cor, 
         method = "ellipse",
         type   = "lower",
         tl.cex = 1.5
         ) 
```

```{r}
#| fig-width: 14
#| fig-height: 14
corrplot(wine.cor, 
         method = "ellipse",
         type   = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.cex = 1.5
         ) 
```

### **Working with mixed layout**

```{r}
#| fig-width: 14
#| fig-height: 14
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")

```

```{r}
#| fig-width: 14
#| fig-height: 14
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

### **Combining corrgram with the significant test**

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)
```

```{r}
#| fig-width: 14
#| fig-height: 14
corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)
```

### **Reorder a corrgram**

```{r}
#| fig-width: 14
#| fig-height: 14
corrplot.mixed(wine.cor, 
               lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               order="AOE",
               tl.col = "black")
```

### **Reordering a correlation matrix using hclust**

```{r}
#| fig-width: 14
#| fig-height: 14
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",
         addrect = 3)
```
